Simple connectome inference from partial correlation statistics in calcium imaging

### 2014-06-26 article.tex
wording

The top of Table \ref{tab:comparison} reports AUROC and AUPRC for all four networks
using in each case partial correlation with different filtering
functions. Except for the last two rows that uses PCA, the exact
inverse of the covariance matrix was used in each case. These results
clearly show the importance of the filters. AUROC increases in average from
0.77 to 0.93. PCA does not really affect AUROC scores but it improves
significantly AUPRC scores. Taking the average over various parameter
setting gives an improvement of 10\% in both AUROC and AUPRC. The last
row shows the final performance of the method specifically tuned for
the challenge (see the Appendix for all details). Although this tuning
was decisive to obtain the best performance at the challenge, it does
not significantly improve neither AUROC nor AUPRC. Note that tuning
was mainly done on the \textit{normal-1} and \textit{normal-4}
datasets, which explains why these two networks are slightly better
retrieved than the others.

\begin{table}[t]
\caption{Evaluation of different methods on the four \textit{normal} datasets\label{tab:comparison}}
\centering
\small
\begin{tabular}{@{}l *{8}{c}@{}}
\hline
  & \multicolumn{4}{c}{AUROC} & \multicolumn{4}{c}{AUPRC} \\
\hline
Methods $\backslash$ normal- & 1 & 2 & 3 & 4 & 1 & 2 & 3 & 4 \\
No  filtering       					& 0.777 & 0.767 & 0.772 & 0.774 & 0.070 & 0.064 & 0.068 & 0.072\\
$ h \circ g \circ f_1$                  & 0.923 & 0.925 & 0.923 & 0.922 & 0.311 & 0.315 & 0.313 & 0.304\\
$ w \circ h \circ g \circ f_1$          & 0.931 & 0.929 & 0.928 & 0.926 & 0.326 & 0.323 & 0.319 & 0.303\\
+ PCA         							& 0.932 & 0.930 & 0.928 & 0.926 & 0.355 & 0.353 & 0.350 & 0.333\\
Averaging           					& 0.938 & 0.936 & 0.936 & 0.932 & 0.391 & 0.389 & 0.385 & 0.374\\
Full method           					& 0.943 & 0.942 & 0.942 & 0.939 & 0.403 & 0.404 & 0.398 & 0.388\\
\hline
Correlation & 0.886 & 0.884 & 0.891 &  0.877 & 0.153 & 0.145 & 0.170 & 0.132\\
GTE & 0.890 & 0.893 & 0.894 & 0.873 & 0.171 & 0.174 & 0.197 & 0.142\\
GENIE3 &\\
\end{tabular}
\caption{TODO \textcolor{red}{version simple!!}}
\label{tab:tab3}
\end{table}

@@

Sections~\ref{sec:filter}~to~\ref{sec:results} describe a simplified version
of the actual winning method. However, one could be interesting in the full
description. This appendix will therefore follow the progress of the main
paper highlighting optimizations.

Let us remark that parameter value choices were made to maximize ROC AUC score
on \textit{normal-1} and, to a lesser extent, \textit{normal-4}.

@@

Theses filters, even if individually they lead to less good performances of the inference algorithm than filters $f_1$ and
$f_2$, the averaging approach over these four low-pass
filters gets better.

As a computation trick, it appears that using a parametrized hard-threshold filter slighlty improves performances. Introducing the parameter $c$ in function $h$ yields to an optimized filter denoted as $h^*$ \begin{align}
h^*(x^t_i) = (x_i^t)^c \text{ } \mathbb{1}(x_i^t \ge \tau) \text{ with } \tau > 0,
\end{align}
where $c$ is set to $0.9$ after experiments.

The magnification can also be optimized by adjusting the effect differently with respect to the global activity. In that sense, we introduce the following function \begin{align}
 w^*(x^{t}_i) &= {(x^{t}_i + 1 )^{\left (1 + \frac{1}{\sum_{j} x^{t}_j}\right )}}^{m}
\end{align}
where $m$ is a piecewise linear function depending on the low-pass filter and on the range of the global activity. We refer to our implementation\footnote{available on \url{https://github.com/asutera/kaggle-connectomics}.} for the values of the parameters and especially the parameter $m$.

@@

In our experiments, we observed that the inference process is quite sensitive
about the choice of the parameter value $\tau$ and the low-pass filter.
Therefore, in order to reduce the variance of the results and thus the
robustness of our approach, we perform an average over a range of parameter
value $\tau$ from $0.100$ to $0.209$. We also perform a weighted average over
the four low-pass filters with weights  $weight_{f_1} = 1$, $weight_{f_2} =
0.9$, $weight_{f_3} = 0.7$ and $weight_{f_4} = 0.01$, tuned to maximize the
ROC AUC on \textit{normal-1}. The weights were adjusted to take into account
that some filters outperform other but there is information hold in every
filter.

@@

The algorithm propose in sections \ref{sec:filter} and \ref{sec:inference} has been optimized at its extreme limit to win the challenge. In this section, we will describe the full tuned algorithm and present some results.

Note that our parameter tuning was mainly
done on the \textit{normal-1} dataset and partly on \textit{normal-4} dataset,
therefore it may justify why these networks are better retrieved despite the
fact that some might be simpler to infer.

% Introduce a custom solution to make the Y matrix asymmetric
% Introduce directivity
\textcolor{red}{Est-ce qu'on ne le ferait pas passer dans l'appendice B? Ou du moins l'équation?}
Since partial correlation is a symmetric measure, the causal mechanism behind the
interaction ($A$ directly causes $B$, $B$ directly causes $A$ or both) can not
be directly inferred\footnote{Also note that two other causal mechanisms might be
implied while having a non-zero partial coefficient: (i) a pair of variables
is induced by a common (hidden) variable; (ii) $A$ (respectively $B$) is
conditionally correlated to a hidden variable affecting $B$ (resp. $A$)
\citep{de2004discovery}.}.
In order to retrieve causal information, we developed an
heuristics based on the following observation. The rise of fluorescence
indicates the activation of a neuron $i$. If a neuron $j$ have also
an increased of fluorescence after a slight delay, this could have been
caused by the neuron $i$. Thus, we count whenever
such events occurred between consecutive time steps
\[
\hat{Z}: z_{ij} = \sum_{t=1}^{T - 1}
    \mathbb{1}((x_j^{t+1} - x_i^t) \in \left[f_1, f_2 \right]) -
    \mathbb{1}((x_i^{t+1} - x_j^t) \in \left[f_1, f_2 \right]), \forall i, j \in V
\]
where $f_1$ and $f_2$ are parameters of the method.
Finally, the weighted average of correlation matrices is combined with $Z$ such
that $Z$ has a contribution of $0.3\%$ to the prediction of an edge.

% For the directory method, only filters defined by Equations
% \ref{eq:low-pass}, \ref{eq:high-pass} and \ref{eq:hard-treshold-filter} were
% used.

>>>>

The top of Table \ref{tab:comparison} reports AUROC and AUPRC for all four networks
using in each case partial correlation with different filtering
functions. Except for the last two rows that uses PCA, the exact
inverse of the covariance matrix was used in each case. These results
clearly show the importance of the filters. AUROC increases in average from
0.77 to 0.93. PCA does not really affect AUROC scores but it improves
significantly AUPRC scores. Taking the average over various parameter
setting gives an improvement of 10\% \textcolor{red}{AS:AUPRC ok but AUROC?} in both AUROC and AUPRC. The last
row shows the final performance of the method specifically tuned for
the challenge (see the Appendix \ref{app:optimized} for all details). Although this tuning
was decisive to obtain the best performance at the challenge, it does
not significantly improve neither AUROC nor AUPRC. Note that tuning
was mainly done on the \textit{normal-1} and \textit{normal-4}
datasets, which explains why these two networks are slightly better
retrieved than the others.

\begin{table}[t]
\caption{Evaluation of different methods on the four \textit{normal} datasets\label{tab:comparison}}
\centering
\small
\begin{tabular}{@{}l *{8}{c}@{}}
\hline
  & \multicolumn{4}{c}{AUROC} & \multicolumn{4}{c}{AUPRC} \\
\hline
Methods $\backslash$ normal- & 1 & 2 & 3 & 4 & 1 & 2 & 3 & 4 \\
No  filtering       					& 0.777 & 0.767 & 0.772 & 0.774 & 0.070 & 0.064 & 0.068 & 0.072\\
$ h \circ g \circ f_1$                  & 0.923 & 0.925 & 0.923 & 0.922 & 0.311 & 0.315 & 0.313 & 0.304\\
$ w \circ h \circ g \circ f_1$          & 0.931 & 0.929 & 0.928 & 0.926 & 0.326 & 0.323 & 0.319 & 0.303\\
+ PCA         							& 0.932 & 0.930 & 0.928 & 0.926 & 0.355 & 0.353 & 0.350 & 0.333\\
Averaging           					& 0.938 & 0.936 & 0.936 & 0.932 & 0.391 & 0.389 & 0.385 & 0.374\\
Full method           					& 0.943 & 0.942 & 0.942 & 0.939 & 0.403 & 0.404 & 0.398 & 0.388\\
\hline
Correlation & 0.886 & 0.884 & 0.891 &  0.877 & 0.153 & 0.145 & 0.170 & 0.132\\
GTE & 0.890 & 0.893 & 0.894 & 0.873 & 0.171 & 0.174 & 0.197 & 0.142\\
GENIE3 &\\
\end{tabular}
\caption{Top: ROC (left) and PR (right) AUC on \text{normal-1,2,3,4} with partial correlation and different filtering functions except the last two rows with an average over various parameter settings.
Bottom: ROC (left) and PR (right) AUC on \text{normal-1,2,3,4} with different methods.}
\label{tab:tab3}
\end{table}

@@

This section is a complete description of the method specifically tuned for the challenge. Basically, it is an optimized version of the method presented in Sections~\ref{sec:filter}~to~\ref{sec:results}, therefore only differences between both version of our method will be highlighted.

Let us remark that tuning
was mainly done on the \textit{normal-1} and \textit{normal-4}
datasets, which explains why these two networks are slightly better
retrieved than the others.

@@

Theses filters lead, taken individually, to worse performances
 than filters $f_1$ and
$f_2$ but the averaging approach over all low-pass
filters gets better \textcolor{red}{AS:Does it necessary to show it numerically?}.

As a computation trick, it appears that using a parametrized hard-threshold filter slightly improves performances. Introducing the parameter $c$ in function $h$ yields to an optimized filter \begin{align}
h^*(x^t_i) = (x_i^t)^c \text{ } \mathbb{1}(x_i^t \ge \tau) \text{ with } \tau > 0,
\end{align}
where $c$ is set to $0.9$ after experiments.

The magnification can also be optimized by adjusting differently the effect depending on the global activity and the low-pass filter used. In that sense, we introduce the following function \begin{align}
 w^*(x^{t}_i) &= {(x^{t}_i + 1 )^{\left (1 + \frac{1}{\sum_{j} x^{t}_j}\right )}}^{m}
\end{align}
where $m$ is a piecewise linear function depending on the low-pass filter and on the range of the global activity. We refer to our implementation\footnote{available on \url{https://github.com/asutera/kaggle-connectomics}.} for the values of the parameters and especially the parameter $m$.

@@

In our experiments, we observed that the inference process is quite sensitive
about the choice of the parameter value $\tau$ and the low-pass filter.
Therefore, in order to reduce the variance of the results and thus the
robustness of our approach, we perform an average over a range of parameter
value $\tau$ from $0.100$ to $0.209$. We also perform a weighted average over
the four low-pass filters with the following parameters

\begin{align*}
weight_{f_1} &= 1,\\
weight_{f_2} &= 0.9, \\
weight_{f_3} &= 0.7, \\ 
weight_{f_4} &= 0.01
\end{align*}

tuned to maximize the
ROC AUC on \textit{normal-1}.

@@

% The algorithm propose in sections \ref{sec:filter} and \ref{sec:inference} has been optimized at its extreme limit to win the challenge. In this section, we will describe the full tuned algorithm and present some results.

% Note that our parameter tuning was mainly
% done on the \textit{normal-1} dataset and partly on \textit{normal-4} dataset,
% therefore it may justify why these networks are better retrieved despite the
% fact that some might be simpler to infer.

% % Introduce a custom solution to make the Y matrix asymmetric
% % Introduce directivity
% \textcolor{red}{Est-ce qu'on ne le ferait pas passer dans l'appendice B? Ou du moins l'équation?}
% Since partial correlation is a symmetric measure, the causal mechanism behind the
% interaction ($A$ directly causes $B$, $B$ directly causes $A$ or both) can not
% be directly inferred\footnote{Also note that two other causal mechanisms might be
% implied while having a non-zero partial coefficient: (i) a pair of variables
% is induced by a common (hidden) variable; (ii) $A$ (respectively $B$) is
% conditionally correlated to a hidden variable affecting $B$ (resp. $A$)
% \citep{de2004discovery}.}.
% In order to retrieve causal information, we developed an
% heuristics based on the following observation. The rise of fluorescence
% indicates the activation of a neuron $i$. If a neuron $j$ have also
% an increased of fluorescence after a slight delay, this could have been
% caused by the neuron $i$. Thus, we count whenever
% such events occurred between consecutive time steps
% \[
% \hat{Z}: z_{ij} = \sum_{t=1}^{T - 1}
%     \mathbb{1}((x_j^{t+1} - x_i^t) \in \left[f_1, f_2 \right]) -
%     \mathbb{1}((x_i^{t+1} - x_j^t) \in \left[f_1, f_2 \right]), \forall i, j \in V
% \]
% where $f_1$ and $f_2$ are parameters of the method.
% Finally, the weighted average of correlation matrices is combined with $Z$ such
% that $Z$ has a contribution of $0.3\%$ to the prediction of an edge.

% % For the directory method, only filters defined by Equations
% % \ref{eq:low-pass}, \ref{eq:high-pass} and \ref{eq:hard-treshold-filter} were
% % used.
